A meta analysis of tournaments and an evaluation of performance in the Iterated Prisoner's Dilemma.

### 2020-01-28 paper/main.tex
address TODO in conclusion
Prepares paper (#13)
* re-run analysis with new data

* clean repo

* move notebooks

* rename files, run black and tweak code

* tests to pytest

* test and implement prepare, process src files

* create clystering analysis script

* update enviroment yml

* add feautures to analysis

* add lines to run on siren

* tweaks to outputs

* add new meausures to analysis

* add probend, noise and turns to cluster analysis

* fix failure

* remove turns as default cluster feauture

* week commit

* write first sections of papers

* fix errors and move tests

* create test

Though only a select pieces of work have been discussed, the IPD literature is
rich, and new strategies and competitions are being published every year~\cite{Glynatsi2019}. The
question, however, still remains the same: what is the best way to play the
game? Compared to other works, whereas a few selected strategies are evaluated
on a small number of tournaments, this manuscript evaluates the performance of \numberofstrategies
strategies in \numberofalltournaments tournaments. These tournaments do not
consist by just standard round robin tournaments, but also by tournaments with noise
and tournaments with a probabilistic ending. The later part of the paper, evaluates
the impact of features on the performance of the strategies using modern
machine learning techniques. These features include measures regarding a
strategy's behaviour and measures regarding the tournaments. The data set used
in this work has been made publicly available~\cite{data} and can be used
for further analysis and insights.

The different tournament types as well as the data collection, which is made
possible due to an open source package called Axelrod-Python,
are covered in Section~\ref{section:data_collection}.
Section~\ref{section:top_performances}, focuses on the best performing
strategies for each type of tournament and overall.
Section~\ref{section:evaluation_of_performance}, explores the traits which
contribute to good performance, and finally the results are summarised in
Section~\ref{section:conclusion}. This manuscripts uses several parameters.
These are introduced in the following sections, however, the full set of
parameters and their definitions are given in Appendix~\ref{app:parameters}.

@@

For the purposes of this manuscript a data set containing results of IPD
tournaments has been generated and is available at~\cite{data}. This was done using the
open source package Axelrod-Python library~\cite{axelrodproject} (APL), and more specifically,
version 3.0.0. APL allows for different types of IPD computer
tournaments to be simulated whilst containing a list of over 180 strategies.
Most of these are strategies described in the literature with a few exceptions
being strategies that have been contributed specifically to the package. This
paper make use of \numberofstrategies strategies implemented in version 3.0.0. A
list of the strategies is given in the Appendix~\ref{app:list_of_players}.
Although APL features several tournament types, this work considers
standard, noisy, probabilistic ending and noisy probabilistic ending
tournaments.

@@

The result summary, Table~\ref{table:output_result}, has \(N\) number of rows
because each row contains information for each strategy that participated in the
tournament. The information includes the strategy's rank, median score, the rate
with which the strategy cooperated $(C_r)$, its match win count and the
probability that the strategy cooperated in the opening move. Moreover, the
probabilities of a strategy being in any of the four states ($CC, CD, DC, DD$),
and the rate of which the strategy cooperated after each state. A feature that
has been manually included is the \textbf{normalised rank}. The rank of a given
strategy, denoted as \(R\), can vary between 0 and \(N-1\). Thus, the normalised rank,
denoted as $r$, is calculated as a strategy's rank divided by the tournament's
size \(N - 1\). In the next section the performance of these strategies is evaluated
based on their normalised rank.

@@

In standard tournaments 10 out of the 15 top strategies are introduced
in~\cite{Harper2017}. These are strategies based on finite state automata (FSM),
hidden markov models (HMM), artificial neural networks (ANN), lookup tables
(LookerUp) and stochastic lookup tables (Gambler) that have been trained using
reinforcement learning algorithms (evolutionary and particle swarm algorithms).
They have been trained to perform well against the strategies
in APL in a standard tournament, thus their performance in the
specific setting was anticipated. DoubleCrosser, and Fool Me Once, are
strategies not from the literature but from the APL. DoubleCrosser
is a strategy that makes use of the number of turns because it is set to defect on
the last two rounds. The strategy was expected to not perform as well in
tournaments where the number of turns is not specified, but the strategy did not
perform well in tournaments with noise either. Finally, Winner
12~\cite{mathieu2017} and DBS~\cite{Au2006} are both from the the literature.
DBS is strategy specifically designed for noisy environments, however, it ranks
highly only in standard ones.

@@

The top strategies in noisy tournaments are shown in Figure~\ref{fig:noisy_results}. These include deterministic strategies, such
as Tit For 2 Tats~\cite{Axelrod1980b}, Slow Tit For Two Tats~\cite{axelrodproject}, Hard Tit For 2 Tats~\cite{Stewart2012}
and Cycler CCCCCD, and strategies which decide their actions based on the
cooperations to defections ratio, such as ShortMem~\cite{Carvalho2013}, Grumpy
and $e$~\cite{axelrodproject}. Slow Tit For Two Tats is the same strategy as 
Tit For 2 Tats, and at the time of writing this manuscript the
contributors of~\cite{axelrodproject} made a new release where the strategy
has been removed. However, for the purpose of this work the strategy is kept.
The Retaliate and
Limited Retaliate strategies are implemented in APL by the
same contributor. They are strategies designed to defect if the opponent has
tricked them more often than \(x\%\) of the times that they have
done the same. Finally, in $4^{\text{th}}$ and $9^{\text{th}}$ place are Hunter
strategies which trying to extort, equivalently, strategies that play cyclically
and defectors.

From Figure~\ref{fig:noisy_results}, it is evident that the normalised rank
distributions in noisy environments are more variant with higher medians
compared to standard tournaments. The distributions are bimodal. This indicates
that although the top ranked strategies mainly performed well, there are several
tournaments that they ranked in the bottom half. The bimodality of the \(r\)
distributions can be explained by Figure~\ref{fig:noisy_results} which gives the
\(r\) distributions for the top 6 strategies over the noise probability \(p_n\).
Note that for \(p_n=0.5\) a strategy corresponds to a random player and for
\(p_n=1\) a strategy is behaving in the exact opposite way than its design.

@@

From Figure~\ref{fig:effect_of_noise} it is evident that the strategies highly
ranked in noisy environments did so because of their performance in tournaments
with \(p_n>0.5\). In tournaments with a noise probability lower than \(0.5\) the
strategies performed poorly which is why their \(r\) distributions are bimodal.
If during the data collection a \(p_n\) strictly less 0.5 was considered then the top ranked strategies would be
different. There are a total of 5661 trials where \(p_n<0.5\) and the top ranked
strategies are given in Table~\ref{table:subset_noise_result}. The median ranks
are lower than before and the top spots are mainly overtaken by Meta strategies
which include NMWE deterministic and NMWE Long Memory. The Meta
strategies~\cite{axelrodproject} create a team of strategies for themselves and
choose to play as a member of their team based on their scores against a given
opponent.

@@

This section presented the winning strategies in a series of IPD tournaments. In
standard tournaments the top spots were dominated by complex strategies that had
been trained using reinforcement learning techniques. In noisy environments,
whether the number of turns was fixed or not, the winning strategies were
deterministic strategies designed to defect if the opponent tricked them more
than a current amount of the times that they had tricked their opponent. However,
if a value of noise strictly less than 0.5 was considered, then the successful
strategies were strategies based on the behaviour of many strategies. In
probabilistic ending tournaments most of the highly ranked strategies were defecting
strategies and trained finite state automata, all by the authors of~\cite{Ashlock2006, Ashlock2014}.
These strategies ranked high due to their performance in tournaments where the probability of the game ending after
each turn was bigger than 0.1.
Finally the performance of all \numberofstrategies
strategies over the \numberofalltournaments tournaments in this manuscript was
assessed on \(\bar{r}\). The top ranked strategies were a mixture of behaviours
that did well in standard tournaments and tournaments with noise, as well as a
few strategies based on teams.

@@

The aim of this section is to explore the features that contribute to a
strategy's successful performance. The features explored are measures regarding a
strategy's behaviour, along with measures regarding the tournaments the
strategies competed in. These are given in Table~\ref{table:manual_features}.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[h]
    \begin{center}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{gcgcgc}
    \toprule
    feature & feature explanation &  source & value type & min value & max value \\
    \midrule
stochastic  &  If a strategy is stochastic & strategy classifier from APL & boolean  & Na &  Na \\
makes use of game &  If a strategy makes used of the game information & strategy classifier from APL & boolean  & Na &  Na \\
makes use of length &  If a strategy makes used of the number of turns & strategy classifier from APL & boolean  & Na &  Na \\
memory usage &  The memory size of a strategy divided by the number of turns & memory size from APL & float & 0 &  1 \\
SSE & A measure of how far a strategy is from ZD behaviour & method described in~\cite{Knight2019} & float & 0 & 1 \\
max cooperating rate $(C_{\text{max}})$  & The biggest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
min cooperating rate $(C_{\text{min}})$ & The smallest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
median cooperating rate $(C_{\text{median}})$ & The median cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
mean cooperating rate $(C_{\text{mean}})$ & The mean cooperating rate in a given tournament  & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{max}}$ & A strategy's cooperating rate divided by the maximum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{min}}$ & A strategy's cooperating rate divided by the minimum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{median}}$ & A strategy's cooperating rate divided by the median  & result summary  & float & 0 & 1\\
$C_r$ / $C_{\text{mean}}$ & A strategy's cooperating rate divided by the mean & result summary  & float & 0 & 1 \\
$C_r$ & The cooperating ratio of a strategy & result summary  & float & 0 & 1 \\
$CC$ to $C$ rate & The probability a strategy will cooperate after a mutual cooperation & result summary  & float & 0 & 1\\
$CD$ to $C$ rate & The probability a strategy will cooperate after being betrayed by the opponent & result summary  & float & 0 & 1 \\
$DC$ to $C$ rate & The probability a strategy will cooperate after betraying the opponent & result summary  & float & 0 & 1 \\
$DD$ to $C$ rate & The probability a strategy will cooperate after a mutual defection & result summary  & float & 0 & 1 \\
$p_n$ & The probability of a player's action being flip at each interaction & trial summary & float & 0 & 1 \\
$n$ & The number of turns & trial summary & integer & 1 & 200 \\
$p_e$ & The probability of a match ending in the next turn & trial summary & float & 0 & 1 \\
$N$ & The number of strategies in the tournament & trial summary & integer & 3 & 195 \\
$k$ & The number of repetitions of a given tournament & trial summary & integer & 10 & 100 \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{The features which are included in the performance evaluation analysis.}
    \label{table:manual_features}
\end{table}

APL makes use of classifiers to classify strategies according to
various dimensions. These determine whether
a strategy is stochastic or deterministic, whether it makes use of the number of
turns or the game's payoffs. The memory usage feature is calculated as the
memory size of strategy (which is specified in the strategies implementation
in the APL) divide by the number of turns. For example, Evolved
FSM 16 Noise 05 has a memory size of 16 and participated in a tournament where
$n$ was 134. In the given tournament Evolved FSM 16 Noise 05 has a memory usage
of 0.119. For tournaments with a probabilistic ending the number of turns was
not collected, so the memory usage feature is not used for probabilistic ending
tournaments. The SSE is a feature introduced in~\cite{Knight2019} which shows
how close a strategy is to behaving as a ZDs, and subsequently, in an
extortionate way. The method identifies the ZDs closest to a given strategy and
calculates the algebraic distance between them, defined as SSE. A SSE value of 1
indicates no extortionate behaviour at all whereas a value of 0 indicates that a
strategy is behaving a ZDs. The rest of the features considered are the $CC$ to
$C$, $CD$ to $C$, $DC$ to $C$, and $DD$ to $C$ rates as well as cooperating
ratio of a strategy. The minimum, maximum, medium and median cooperating ratios
of each tournament are also included, and finally the number of turns, the
number of strategies, the number of repetitions and the probabilities of noise
and the game ending are also included.

Table~\ref{table:correlations} shows the correlation coefficients between the
features of Table~\ref{table:manual_features} the median score and the median
normalised rank. Note that the correlation for the classifiers is
not included because they are binary variables and they will be evaluated using a
different method. The correlation coefficients for all the features in
Table~\ref{table:manual_features} against themselves have also been calculated and a
graphical representation can be found in the Appendix~\ref{app:correlations}.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[!htbp]
    \begin{center}
    \resizebox{.9\textwidth}{!}{
        \input{correlations_table.tex}
    }
\end{center}
\caption{Correlations table between the features of Table~\ref{table:manual_features}
the normalised rank and the median score.}\label{table:correlations}
\end{table}

In standard tournaments the features  $CC$ to $C$, $C_r$, $C_r / C_{\text{max}}$
and the cooperating ratio compared to $C_{\text{median}}$ and $C_{\text{mean}}$
have a moderate negative effect on the normalised rank, and a moderate positive
on the median score. The SSE error and the $DD$ to $C$ have the opposite
effects. Thus, in standard tournaments behaving cooperatively corresponds to a
more successful performance. Even though being nice pays off,
that's not true against defective strategies. Cooperating after a mutual
defection lowers a strategy's success.
Figure~\ref{fig:rates_of_winners_in_standard_tournaments} confirms that the
winners of standard tournaments always cooperate after a mutual cooperation and
almost always defects after a mutual defection.

@@

Compared to standard tournaments, in both noisy and in probabilistic ending
tournaments the higher the rates of cooperation the lower a strategy's success
and median score. A strategy would want to cooperate less than both
the mean and median cooperator in such settings. In probabilistic ending
tournaments the correlations coefficients have a larger values, indicating a
stronger effect. Thus a strategy will be punished more by it's cooperative
behaviour in probabilistic ending environments, this was seen in Section~\ref{section:evaluation_of_performance}
as well. The distributions of the $C_r$ of the winners in
both tournaments is given by Figure~\ref{fig:c_r_distributions}. It confirms
that the winners in noisy tournaments cooperated less than 35\% of the times
and in probabilistic ending tournaments less than 10\%.
In noisy probabilistic ending tournaments and in over all the tournaments' results,
the only features that had a moderate affect are $C_r/C_{\text{mean}},
C_r/C_{\text{max}}$ and $C_r$. In such environments cooperative behaviour
appears to be punished by not as much as in noisy and probabilistic ending
tournaments.

@@

To further explore the features that contribute to a strategy's success the
performances are divided into clusters based on whether they were successful or
not. A random forest approach~\cite{breiman2001} is then applied to each
performance to predict the cluster to which it has been assigned to. This allows
for the importance of each feature in the classification to be calculated. In
essence, to calculate which are the important features when one is trying to
predict whether a performance was successful or not.

The performances are clustered into successful and unsuccessful clusters based on
4 different approaches. More specifically:

\begin{itemize}
    \item \textbf{Approach 1:} The performances are divided into two clusters based
    on whether their performance was in the top \(5\%\) of their respective tournaments.
    Thus, whether \(r\) was smaller or larger than 0.05.
    \item \textbf{Approach 2:} The performances are divided into two clusters based
    on whether their performance was in the top \(25\%\) of their respective tournaments.
    Thus, whether \(r\) was smaller or larger than 0.25.
    \item \textbf{Approach 3:} The performances are divided into two clusters based
    on whether their performance was in the top \(50\%\) of their respective tournaments.
    Thus, whether \(r\) was smaller or larger than 0.50.
    \item \textbf{Approach 4:} The performances are clustered based on their normalised rank and
    their median score by a \(k-\)means algorithm~\cite{Arthur2007}. The number of
    clusters is not deterministically chosen but it is based on the silhouette
    coefficients~\cite{Rousseeuw1987}.
\end{itemize}

The random forest method constructs many individual decision trees and the
predictions from all trees are pooled to make the final prediction. The random
forest models are trained on a training set of 70\% of the tournaments results.
The accuracy of each model based on $R^2$ and the number of clusters for each
tournament type (because in the case of Approach 4 it is not
deterministically chosen) are given by Table~\ref{table:accuracy_random_forest}.
The out of the bag error (OOB)~\cite{hastie2005} has also been calculated. The
models fit well, and a high value of both the accuracy measures on the test data
and the OOB error indicate that the model is not over fitting.

>>>>

Though only a select pieces of work have been discussed, the IPD literature is
rich, and new strategies and competitions are being published every year. The
question, however, still remains the same: what is the best way to play the
game? Compared to other works, whereas a few selected strategies are evaluated
on a small number of tournaments, this manuscript evaluates the performance of \numberofstrategies
strategies in \numberofalltournaments tournaments. These tournaments do not
consist by just standard round robin tournaments, but also by tournaments with noise
and tournaments with a probabilistic ending. The later part of the paper, evaluates
the impact of features on the performance of the strategies using modern
machine learning techniques. These features include measures regarding a
strategy's behaviour and measures regarding the tournaments. The data set used
in this work has been made publicly available~\cite{data} and can be used
for further analysis and insights.

The different tournament types as well as the data collection, which is made
possible due an open source package called Axelrod-Python~\cite{axelrodproject},
are covered in Section~\ref{section:data_collection}.
Section~\ref{section:top_performances}, focuses on the best performing
strategies for each type of tournament and overall.
Section~\ref{section:evaluation_of_performance}, explores the traits which
contribute to good performance, and finally the results are summarised in
Section~\ref{section:conclusion}. This manuscripts uses several parameters.
These are introduced in the following sections, however, the full set of
parameters and their definitions are given in Appendix~\ref{app:parameters}.

@@

For the purposes of this manuscript a data set containing results of IPD
tournaments has been generated and is available at~\cite{data}. This was done using the
open source package Axelrod-Python~\cite{axelrodproject}, and more specifically,
version 3.0.0. Axelrod-Python allows for different types of IPD computer
tournaments to be simulated whilst containing a list of over 180 strategies.
Most of these are strategies described in the literature with a few exceptions
being strategies that have been contributed specifically to the package. This
paper make use of \numberofstrategies strategies implemented in version 3.0.0. A
list of the strategies is given in the Appendix~\ref{app:list_of_players}.
Though Axelrod-Python features several tournament types, this work considers
only standard, noisy, probabilistic ending and noisy probabilistic ending
tournaments.

@@

The result summary, Table~\ref{table:output_result}, has \(N\) number of rows
because each row contains information for each strategy that participated in the
tournament. The information includes the strategy's rank, median score, the rate
with which the strategy cooperated $(C_r)$, its match win count and the
probability that the strategy cooperated in the opening move. Moreover, the
probabilities of a strategy being in any of the four states ($CC, CD, DC, DD$),
and the rate of which the strategy cooperated after each state. A measure that
has been manually included is the \textbf{normalised rank}. The normalised rank,
denoted as $r$, is calculated as a strategy's rank divided by the tournament's
size ($N$). In the next section the performance of these strategies is evaluated
based on their normalised rank.

@@

In standard tournaments 10 out of the 15 top strategies are introduced
in~\cite{Harper2017}. These are strategies based on finite state automata (FSM),
hidden markov models (HMM), artificial neural networks (ANN), lookup tables
(LookerUp) and stochastic lookup tables (Gambler) that have been trained using
reinforcement learning algorithms (evolutionary and particle swarm algorithms).
They have been trained to perform well against the strategies
in~\cite{axelrodproject} in a standard tournament, thus their performance in the
specific setting was anticipated. DoubleCrosser, and Fool Me Once, are
strategies not from the literature but from~\cite{axelrodproject}. DoubleCrosser
is a strategy that makes use of the number of turns because is set to defect on
the last two rounds. The strategy was expected to not perform as well in
tournaments where the number of turns is not specified, but the strategy did not
perform well in tournaments with noise either. Finally, Winner
12~\cite{mathieu2017} and DBS~\cite{Au2006} are both from the the literature.
DBS is strategy specifically designed for noisy environments, however, it ranks
highly only in standard ones.

@@

The top strategies in noisy tournaments are shown in Figure~\ref{fig:noisy_results}. These include deterministic strategies, such
as Tit For 2 Tats~\cite{Axelrod1980b}, Slow Tit For Two Tats~\cite{axelrodproject}, Hard Tit For 2 Tats~\cite{Stewart2012}
and Cycler CCCCCD, and strategies which decide their actions based on the
cooperations to defections ratio, such as ShortMem~\cite{Carvalho2013}, Grumpy
and $e$~\cite{axelrodproject}. Slow Tit For Two Tats is the same strategy as 
Tit For 2 Tats, and at the time of writing this manuscript the
contributors of~\cite{axelrodproject} made a new release where the strategy
has been removed. However, for the purpose of this work the strategy is kept.
The Retaliate and
Limited Retaliate strategies are implemented in~\cite{axelrodproject} by the
same contributor. They are strategies designed to defect if the opponent has
tricked them more often than \(x\%\) of the times that they have
done the same. Finally, in $4^{\text{th}}$ and $9^{\text{th}}$ place are Hunter
strategies which trying to extort, equivalently, strategies that play cyclically
and defectors.

From Figure~\ref{fig:noisy_results}, it is evident that the normalised rank
distributions in noisy environments are more variant with higher medians
compared to standard tournaments. The distributions are bimodal. This indicates
that although the top ranked strategies mainly performed well, there are several
tournaments that they ranked in the bottom half. To gain a better understanding
of the behaviour in noisy tournaments, the \(r\) distributions for the top 6
of Figure~\ref{fig:noisy_results} strategies over the noise probability \(p_n\),
are given in Figure~\ref{fig:effect_of_noise}.

@@

Figure~\ref{fig:effect_of_noise} shows that for \(p_n\) values lower than 0.5
Grumpy, Tit For 2 Tats and Slow Tit For Two Tat perform moderately, and \(e\),
Cycle Hunter and Ricky QLearner perform poorly. At \(p_n=0.5\) all the
distributions become bimodal. This is because with a noise probability of 0.5,
all strategies correspond to a random player. Interestedly, for a \(p_n\) larger
than 0.5 all of the 6 strategies become successful. Note that a value \(p_n=1\)
corresponds to a strategy playing opposite from what it intended to. Thus, it is
demonstrated that the successful strategies is noisy tournaments are sometimes
effective when \(p_n=0.5\) but overall they are very successful whn they are
playing opposite from their original design. If during the data collection a
\(p_n\) strictly less 0.5 was considered then the top ranked strategies would be
different. There are a total of 5661 trials where \(p_n<0.5\) and the top ranked
strategies are given in Table~\ref{table:subset_noise_result}. The median ranks
are lower than before and the top spots are mainly overtaken by Meta strategies
which include NMWE deterministic and NMWE Long Memory. The Meta
strategies~\cite{axelrodproject} create a team of strategies for themselves and
choose to play as a member of their team based on their scores against a given
opponent.

@@

This section presented the winning strategies in a series of IPD tournaments. In
standard tournaments the top spots were dominated by complex strategies that had
been trained using reinforcement learning techniques. In noisy environments,
whether the number of turns was fixed or not, the winning strategies were
deterministic strategies designed to defect if the opponent tricked them more
than a current amount of the times that they had tricked their opponent. However,
if a value of noise strictly less than 0.5 was considered, then the successful
strategies were strategies based on teams. In
probabilistic ending tournaments most of the winning strategies were defecting
strategies and trained finite state automata, designed by the same authors.
These strategies only did better when the probability of the game ending after
each turn was increased.
Finally the performance of all \numberofstrategies
strategies over the \numberofalltournaments tournaments in this manuscript was
assessed on \(\bar{r}\). The top ranked strategies were a mixture of behaviours
that did well in standard tournaments and tournaments with noise, as well as a
few strategies based on teams.

@@

The aim of this section is to explore the factors that contribute to a
strategy's successful performance. The factors explored are measures regarding a
strategy's behaviour, along with measures regarding the tournaments the
strategies competed in. These are given in Table~\ref{table:manual_measures}.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[h]
    \begin{center}
    \resizebox{\textwidth}{!}{
    \begin{tabular}{gcgcgc}
    \toprule
    measure & measure explanation &  source & value type & min value & max value \\
    \midrule
stochastic  &  If a strategy is stochastic & strategy classifier from~\cite{axelrodproject} & boolean  & Na &  Na \\
makes use of game &  If a strategy makes used of the game information & strategy classifier from~\cite{axelrodproject} & boolean  & Na &  Na \\
makes use of length &  If a strategy makes used of the number of turns & strategy classifier from~\cite{axelrodproject} & boolean  & Na &  Na \\
memory usage &  The memory size of a strategy divided by the number of turns & memory size from~\cite{axelrodproject} & float & 0 &  1 \\
SSE & A measure of how far a strategy is from ZD behaviour & method described in~\cite{Knight2019} & float & 0 & 1 \\
max cooperating rate $(C_{\text{max}})$  & The biggest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
min cooperating rate $(C_{\text{min}})$ & The smallest cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
median cooperating rate $(C_{\text{median}})$ & The median cooperating rate in a given tournament  & result summary  & float & 0 & 1\\
mean cooperating rate $(C_{\text{mean}})$ & The mean cooperating rate in a given tournament  & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{max}}$ & A strategy's cooperating rate divided by the maximum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{min}}$ & A strategy's cooperating rate divided by the minimum & result summary  & float & 0 & 1 \\
$C_r$ / $C_{\text{median}}$ & A strategy's cooperating rate divided by the median  & result summary  & float & 0 & 1\\
$C_r$ / $C_{\text{mean}}$ & A strategy's cooperating rate divided by the mean & result summary  & float & 0 & 1 \\
$C_r$ & The cooperating ratio of a strategy & result summary  & float & 0 & 1 \\
$CC$ to $C$ rate & The probability a strategy will cooperate after a mutual cooperation & result summary  & float & 0 & 1\\
$CD$ to $C$ rate & The probability a strategy will cooperate after being betrayed by the opponent & result summary  & float & 0 & 1 \\
$DC$ to $C$ rate & The probability a strategy will cooperate after betraying the opponent & result summary  & float & 0 & 1 \\
$DD$ to $C$ rate & The probability a strategy will cooperate after a mutual defection & result summary  & float & 0 & 1 \\
$p_n$ & The probability of a player's action being flip at each interaction & trial summary & float & 0 & 1 \\
$n$ & The number of turns & trial summary & integer & 1 & 200 \\
$p_e$ & The probability of a match ending in the next turn & trial summary & float & 0 & 1 \\
$N$ & The number of strategies in the tournament & trial summary & integer & 3 & 195 \\
$k$ & The number of repetitions of a given tournament & trial summary & integer & 10 & 100 \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{The measures which are included in the performance evaluation analysis.}
    \label{table:manual_measures}
\end{table}

Axelrod-Python makes use of classifiers to classify strategies according to
various dimensions. These determine whether
a strategy is stochastic or deterministic, whether it makes use of the number of
turns or the game's payoffs. The memory usage measure is calculated as the
memory size of strategy (which is specified in the strategies implementation
in~\cite{axelrodproject}) divide by the number of turns. For example, Evolved
FSM 16 Noise 05 has a memory size of 16 and participated in a tournament where
$n$ was 134. In the given tournament Evolved FSM 16 Noise 05 has a memory usage
of 0.119. For tournaments with a probabilistic ending the number of turns was
not collected, so the memory usage measure is not used for probabilistic ending
tournaments. The SSE is a measure introduced in~\cite{Knight2019} which shows
how close a strategy is to behaving as a ZDs, and subsequently, in an
extortionate way. The method identifies the ZDs closest to a given strategy and
calculates the algebraic distance between them, defined as SSE. A SSE value of 1
indicates no extortionate behaviour at all whereas a value of 0 indicates that a
strategy is behaving a ZDs. The rest of the factors considered are the $CC$ to
$C$, $CD$ to $C$, $DC$ to $C$, and $DD$ to $C$ rates as well as cooperating
ratio of a strategy. The minimum, maximum, medium and median cooperating ratios
of each tournament are also included, and finally the number of turns, the
number of strategies, the number of repetitions and the probabilities of noise
and the game ending are also included.

Table~\ref{table:correlations} shows the correlation coefficients between the
measures of Table~\ref{table:manual_measures} the median score and the median
normalised rank. Note that the correlation for the classifiers is
not included because they are binary variables and they will be evaluated using a
different method. The correlation coefficients for all the measures in
Table~\ref{table:manual_measures} against themselves have also been calculated and a
graphical representation can be found in the Appendix~\ref{app:correlations}.

\newcolumntype{g}{>{\columncolor{Gray}}c}
\begin{table}[!htbp]
    \begin{center}
    \resizebox{.9\textwidth}{!}{
        \input{correlations_table.tex}
    }
\end{center}
\caption{Correlations table between the measures of Table~\ref{table:manual_measures}
the normalised rank and the median score.}\label{table:correlations}
\end{table}

In standard tournaments the measures  $CC$ to $C$, $C_r$, $C_r / C_{\text{max}}$
and the cooperating ratio compared to $C_{\text{median}}$ and $C_{\text{mean}}$
have a moderate negative effect on the normalised rank, and a moderate positive
on the median score. The SSE error and the $DD$ to $C$ have the opposite
effects. Thus, in standard tournaments behaving cooperatively corresponds to a
more successful performance. Even though being nice pays off,
that's not true against defective strategies. Cooperating after a mutual
defection lowers a strategy's success.
Figure~\ref{fig:rates_of_winners_in_standard_tournaments} confirms that the
winners of standard tournaments always cooperate after a mutual cooperation and
almost always defects after a mutual defection.

@@

Compared to standard tournaments, in both noisy and in probabilistic ending
tournaments the higher the rates of cooperation the lower a strategy's success
and median score. A strategy would want to cooperate less than both
the mean and median cooperator in such settings. In probabilistic ending
tournaments the correlations coefficients have a larger values, indicating a
stronger effect. Thus a strategy will be punished more by it's cooperative
behaviour in probabilistic ending environments, this was seen in Section~\ref{section:evaluation_of_performance}
as well. The distributions of the $C_r$ of the winners in
both tournaments is given by Figure~\ref{fig:c_r_distributions}. It confirms
that the winners in noisy tournaments cooperated less than 35\% of the times
and in probabilistic ending tournaments less than 10\%.
In noisy probabilistic ending tournaments and in over all the tournaments' results,
the only measures that had a moderate affect are $C_r/C_{\text{mean}},
C_r/C_{\text{max}}$ and $C_r$. In such environments cooperative behaviour
appears to be punished by not as much as in noisy and probabilistic ending
tournaments.

@@

The performances are clustered based on the normalised rank. More specifically,
they are clustered 3 times into 2 different clusters based on  on whether their
normalised rank was in the top 5\%, 25\% or 50\% respectively. A random forest
approach~\cite{breiman2001} is then applied to each performance to predict the cluster to
which it has been assigned to. The random forest method
constructs many individual decision trees and the predictions from all trees are
pooled to make the final prediction. The random forest models are trained on a
training set of 70\% of the tournaments results. The accuracy of each model
based on $R^2$ are given by Table~\ref{table:accuracy_random_forest}. The out of
the bag error~\cite{hastie2005} has also been calculated. The models fit well,
and a high value of both the accuracy measure on the test data and the OOB error
indicate that the model is not over fitting.

The performances have also been clustered based on their normalised rank and
their median score by a \(k-\)means algorithm~\cite{Arthur2007}. The number of
clusters is not deterministically chosen but it is based on the silhouette
coefficients~\cite{Rousseeuw1987}. The chosen cluster for each tournament type,
as well as the accuracy for random forest models are also given in
Table~\ref{table:accuracy_random_forest}.

\begin{table}[!htbp]
    \begin{center}
        \resizebox{.9\textwidth}{!}{
        \begin{tabular}{lccccc}
    \toprule
    Tournament type & Clustering on & Number of clusters & $R^2$ training data &  $R^2$ test data  & $R^2$ OOB score\\
    \midrule
    standard  & top 5\% $r$              & 2  & 0.998831  & 0.987041 & 0.983708 \\
              & top 25\% $r$             & 2  & 0.998643  & 0.978626 & 0.969202 \\
              & top 50\% $r$             & 2  & 0.998417  & 0.985217 & 0.976538 \\
              & $r$  \& normalised score & 2  & 0.998794  & 0.990677 & 0.982959 \\
    \midrule
    noisy     & top 5\% $r$              & 2 & 0.996677  & 0.950572 & 0.935383\\
              & top 25\% $r$             & 2 & 0.996677  & 0.950572 & 0.935383\\
              & top 50\% $r$             & 2 & 0.996677  & 0.950572 & 0.935383\\
              & $r$  \& normalised score & 3 & 0.996677  & 0.950572 & 0.935383\\
    \midrule
    probabilistic ending & top 5\% $r$              & 2 & 0.999592  & 0.995128 & 0.992819 \\
                         & top 25\% $r$             & 2 & 0.999592  & 0.995128 & 0.992819 \\
                         & top 50\% $r$             & 2 & 0.999592  & 0.995128 & 0.992819 \\
                         & $r$  \& normalised score & 2 & 0.999592  & 0.995128 & 0.992819 \\
    \midrule
    noisy probabilistic ending & top 5\% $r$              & 2 & 0.990490  & 0.813905 & 0.791418\\
                               & top 25\% $r$             & 2 & 0.990490  & 0.813905 & 0.791418\\
                               & top 50\% $r$             & 2 & 0.990490  & 0.813905 & 0.791418\\
                               & $r$  \& normalised score & 4 & 0.990490  & 0.813905 & 0.791418\\
    \midrule
    over \numberofalltournaments tournaments & top 5\% $r$               & 2 & 0.993396 & 0.913409 & 0.898059 \\
                                             & top 25\% $r$              & 2 & 0.993396 & 0.913409 & 0.898059 \\
                                             & top 50\% $r$              & 2 & 0.993396 & 0.913409 & 0.898059 \\
                                             & $r$  \& normalised score  & 3 & 0.993396 & 0.913409 & 0.898059 \\
    \bottomrule
        \end{tabular}}
    \end{center}
    \caption{Accuracy metrics for random forest models.}
    \label{table:accuracy_random_forest}
\end{table}

The importance that the measures of Table~\ref{table:manual_measures} had on
each classification task; to which cluster a performance was assigned to based
on the normalised rank, and their normalised rank and median score have been
calculated and are given by Figures~\ref{fig:clustering_importance_standard},
\ref{fig:clustering_importance_noise},~\ref{fig:clustering_importance_probend},
\ref{fig:clustering_importance_probend_noise}
and~\ref{fig:clustering_importance_overall}. These show that the classifiers
stochastic, make use of game and make use of length have no significant effect,
and several of the measures that are highted by the importance are inline with
the correlation results. Moreover, the smoothing parameter \(k\) appears to no
have a significant effect either. The most important measures based on the
random forest analysis were $C_{r} / C_{median}$ and $C_r / C_{mean}$.
